Site definition possibilites: Features excavated in a transect

### 2018-03-18 analysis/paper/paper.Rmd
unnecessaries deleted.

Thank you for the invitation to speak in this session. My name is Sophie Schmidt, I am research assistant in Cologne.

@@

table3 <- siedlung_BefSchnitt %>% group_by(kultur) %>% arrange(kultur) %>% summarise(count = n())

@@

There are 4770 features catalogued as belonging to these periods or as undated and you can see the specific number of settlement features belonging to the periods in this table. It is easy to see, that only with the beginning of the late Neolithic enough features for statistical analysis were excavated.

Because of the test trenching I assume that all features and sites along the transect have been found, even though there are gaps between the actual excavation areas. Machine trenching has been shown to be the most effective way to find sites [@hey_2006]. Admittedly even then not all sites are discovered. Especially neolithic ones have the least probability to turn up, which may be one reason for the small number of middle neolithic features in the transect.

@@

The term "site" has been at the center of a lively debate in survey archaeology over the last 40 years. Until the 70's it has mostly been defined as a place - a findspot, which is similar to German archaeology, where the term used, "Fundplatz", means exactly that. In the German archaeological inventories, though, a site may contain several culturally distinct loci "Fundstellen", which are ususally catalogued as a settlement, a burial ground, a hoard or as a single find spot of a period.

@@

The **Early Bronze Age or Unetice Culture** has only 25 features in this analysis, which are all found in the Western part of the transect. There are five houses and it is notable, that the distances the houses have to each other dominate the distribution. The turning point of the cumulative analysis about 220 m, which correlates quite well with the sizes two known Unetice ring ditches in Schloßvippach, Thuringia [@walter_2007] and  Zwenkau-Eytrah, Saxony [@huth_1998]. Ditches are often interpreted as settlement boundaries.
The distance between settlements which I propose here has an equivalent (430 m) at the excavation of the same street B6n further west, between Benzigerode and Heimburg, where two early bronze age settlements are cut.

>>>>

@@

@@

There are 'r  #sum(table$count)' 4770 features catalogued as belonging to these periods or as undated and you can see the specific number of settlement features belonging to the periods in this table. It is easy to see, that only with the beginning of the late Neolithic enough features for statistical analysis were excavated.

Because of the test trenching I assume that all features and sites along the transect have been found, even though there are gaps between the actual excavation areas. Machine trenching has been shown to be the most effective way to find sites [@hey_2006]. Admittedly even then not all sites are discovered. Especially neolithic ones have the least probability to turn up, which might be one reason for the small number of middle neolithic features in the transect.

@@

The term "site" has been at the center of a lively debate in survey archaeology over the last 40 years. Until the 70's it has mostly been defined as a place - a findspot, which is similar to German archaeology, where the term used, "Fundplatz", means exactly that. In the German archaeological inventories, a site may contain several culturally distinct loci "Fundstellen", which are ususally catalogued as a settlement, a burial ground, a hoard or as a single find spot of a period.

@@

The **Early Bronze Age or Unetice Culture** has only 25 features in this analysis, which are all found in the Western part of the transect. There are five houses and it is notable, that the distances the houses have to each other dominate the distribution. The turning point of the cumulative analysis about 220 m, which correlates quite well with the sizes of known Unetice ring ditches in Schloßvippach, Thuringia [@walter_2007] and  Zwenkau-Eytrah, Saxony [@huth_1998]. Ditches are often interpreted as settlement boundaries.
The distance between settlements which I propose here has an equivalent (430 m) at the excavation of the same street B6n further west, between Benzigerode and Heimburg, where two early bronze age settlements are cut.
